---
title: "To Do Machine Learning Course"
author: "Jan-Philipp Kolb"
date: "12 April 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Parts not used yet

- The datasets we use for this course
- What to do with missing data?

## The package `rpart` 

- `rpart` - recursive partitioning and regression trees

### The `kyphosis` dataset 

- Data on Children who have had Corrective Spinal Surgery
- The German word is Rundrücken

```{r,eval=F}
?kyphosis
```

<!--
vertebrae - Wirbel
-->


- Number - the number of vertebrae involved
- Start - the number of the first (topmost) vertebra operated on.

```{r}
kable(head(kyphosis))
```


## [Classification Tree example](https://www.statmethods.net/advstats/cart.html)

<!--
https://www.statmethods.net/advstats/cart.html
-->

```{r}
fit <- rpart(Kyphosis ~ Age + Number + Start,
   method="class", data=kyphosis)
```

```{r}
printcp(fit) # display the results 
```


```{r}
plotcp(fit) 
```


```{r}
summary(fit)
```

```{r}
# plot tree 
plot(fit, uniform=TRUE, 
   main="Classification Tree for Kyphosis")
text(fit, use.n=TRUE, all=TRUE, cex=.8)
```

```{r,eval=F}
# create attractive postscript plot of tree 
post(fit, file = "D:/Daten/GitHub/machine_learning/slides/figure/tree.ps", 
   title = "Classification Tree for Kyphosis")
```

