---
title: "Brainstorming - find topics"
author: "Jan-Philipp Kolb"
date: "15 Januar 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## [Supervised vs unsupervised learning](https://towardsdatascience.com/supervised-vs-unsupervised-learning-14f68e32ea8d) (essentially regression/classification vs EDA)


- [Supervised learning](https://www.datasciencecentral.com/profiles/blogs/supervised-learning-everything-you-need-to-know)
- Exploratory data analysis



Bias-variance tradeoff

Curse of dimensionality

Cross-validation +++

## LASSO, Ridge and elasticnet

- [what is better in which situation](https://web.stanford.edu/~hastie/StatLearnSparsity_files/SLS_corrected_1.4.16.pdf)

Kernal regression? +++

- Bagging and boosting 


CART (classification and regression trees)

Classification

Support vector machines (both classification and regression). +++

Neural networks (plus gradient descent)
       Perceptron
       Multilayer perceptron (aka  feedforward neural network with two or more layers)
       backpropagation

Causal graphs/search algorithms


<!--
https://www.r-bloggers.com/automated-dashboard-for-classification-neural-network-in-r/

-->


## Further ideas

- Present the usage of databases like PostgreSQL or MySQL
- Present possibilities like AWS






## First part

- importing data

- what to do with missing data
- introduce/explain datasets used (OSM dataset, PUF GESIS Panel)
- parallelization

## Second part - lasso regression

- cross validation
- evaluation (confusion matrix)
- glmnet package


## Third part

- gbm package
- ensemble methods - gradient boosting

(caret package)

## Sixth Part

[](http://blog.conceptnet.io/posts/2017/how-to-make-a-racist-ai-without-really-trying/)


## 

[HDBSCAN](https://cran.r-project.org/web/packages/dbscan/vignettes/hdbscan.html)

<!--
better implemented in python
-->

- 


- how to run python from R

```{r}
library(reticulate)
```


[hdbscan](https://pypi.org/project/hdbscan/) - slow in R

you should use instead of kmeans


dimension reduction procedure

[Umap](https://github.com/lmcinnes/umap)

https://arxiv.org/pdf/1802.03426.pdf

<!--
Bescheid sagen, dass Python auf den Rechnern installiert sein sollte
-->

- mention interfaces to databases



class myscript{public static void main string[] args{System.out.println("Hello World");}}




