---
title: "Machine Learning - Solution/Exercises"
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output:
  beamer_presentation: 
    colortheme: dolphin
    fig_height: 3
    fig_width: 5
    fig_caption: no
    fonttheme: structuresmallcapsserif
    highlight: haddock
    theme: Dresden
  pdf_document: 
    keep_tex: yes
    toc: yes
  slidy_presentation: 
    css: mycss.css
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## [Exercise - `rpart` Kyphosis](https://www.r-exercises.com/2016/12/13/recursive-partitioning-and-regression-trees-exercises/)

Consider the Kyphosis data frame(type `help("kyphosis")` for more details), that contains:

- Kyphosis: a factor with levels absent present indicating if a kyphosis (a type of deformation) was present after the operation.
- Age: in months.
- Number: the number of vertebrae involved.
- Start: the number of the first (topmost) vertebra operated on.

1) Build a tree to classify Kyphosis from Age, Number and Start.

### Consider the tree build above.

1) Which variables are used to explain Kyphosis presence?
2) How many observations contain the terminal nodes.

## [Solution - `rpart` Kyphosis](https://www.r-exercises.com/2016/12/13/recursive-partitioning-and-regression-trees-solutions/)

```{r}
library('rpart')
TREE <- rpart(Kyphosis ~ Age + Number + Start, 
              data=kyphosis,method="class")
TREE
```

```{r}
plot(TREE)
text(TREE,use.n=T)
```

## Answers

1) Which variables are used to explain Kyphosis presence?

- The variables are Start and Age

2) How many observations contain the terminal nodes.

- *denotes terminal nodes. Each nodes has 29, 12, 14, 7 and 19 observations