---
title: "ML Exercises - Gradient Boosting"
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output: 
  beamer_presentation: 
    theme: Dresden
    colortheme: dolphin
    fig_height: 3
    fig_width: 5
    fig_caption: no
    fonttheme: structuresmallcapsserif
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=T,message=F,warning=F)
```

## [eXtremely Boost your machine learning Exercises (Part-1)](https://www.r-exercises.com/2017/09/24/extremely-boost-your-machine-learning-exercises-part-1/)

```{r,eval=F,echo=F}
install.packages("xgboost")
```


- eXtreme Gradient Boosting is a machine learning model which became really popular few years ago after winning several Kaggle competitions. 
- It is very powerful algorithm that use an ensemble of weak learners to obtain a strong learner. 
- Its R implementation is available in xgboost package and it is really worth including into anyone’s machine learning portfolio.

<!--
This is the first part of eXtremely Boost your machine learning series. For other parts follow the tag xgboost.

Answers to the exercises are available here.

If you obtained a different (correct) answer than those listed on the solutions page, please feel free to post your answer as a comment on that page.
-->

## Boosting Exercises - first part

### Exercise 1
Load `xgboost` library and download German Credit dataset. Your goal will be to predict creditability (the first column in the dataset).

### Exercise 2
Convert columns `c(2,4,5,7,8,9,10,11,12,13,15,16,17,18,19,20)` to factors and then encode them as dummy variables. HINT: use the command `model.matrix()`

### Exercise 3
Split data into training and test set 700:300. Create `xgb.DMatrix` for both sets with Creditability as label.

## Boosting Exercises - second part

### Exercise 4
Train `xgboost` with logistic objective and 30 rounds of training and maximal depth 2.

### Exercise 5
To check model performance calculate test set classification error.

### Exercise 6
Plot predictors importance.

## Boosting Exercises - third part

### Exercise 7
Use `xgb.train()` instead of `xgboost()` to add both train and test sets as a watchlist. Train model with same parameters, but 100 rounds to see how it performs during training.

### Exercise 8
Train model again adding AUC and Log Loss as evaluation metrices.

### Exercise 9
Plot how AUC and Log Loss for train and test sets was changing during training process. Use plotting function/library of your choice.

### Exercise 10
Check how setting parameter eta to 0.01 influences the AUC and Log Loss curves.
image_pdf

## [Solutions: boosting exercises](https://www.r-exercises.com/2017/09/24/extremely-boost-your-machine-learning-solutions-part-1/)

### Solution Exercise 1 - import dataset

```{r}
library(xgboost)
```


```{r,eval=F}
url <- "http://freakonometrics.free.fr/german_credit.csv"
credit <- read.csv(url, header = TRUE, sep = ",")
```


```{r,eval=F,echo=F}
save(credit,file="../data/german_credit.RData")
```

```{r,echo=F}
load("../data/german_credit.RData")
```

```{r}
head(credit)
```

## Solutions boosting exercises - first part

### Solution Exercise 2 - convert columns

```{r}
factor_columns <- c(2,4,5,7,8,9,10,11,12,13,15,16,17,18,19,20)
for(i in factor_columns) credit[,i] <- as.factor(credit[,i])
X <- model.matrix(~ . - Creditability, data=credit)
```

###  Solution  Exercise 3 

```{r}
inTraining <- sample(1:nrow(credit),size=700)
dtrain <- xgboost::xgb.DMatrix(X[inTraining,],
                      label=credit$Creditability[inTraining])
dtest <- xgboost::xgb.DMatrix(X[-inTraining,],
                     label=credit$Creditability[-inTraining])
```

## Solutions boosting exercises - second part

### Solution Exercise 4 - train `xgboost` model 

```{r}
model <- xgboost(data = dtrain,
                 max_depth = 2,
                 nrounds = 30,
                 objective = "binary:logistic")
```

## Solutions boosting exercises - third part

### Solution    Exercise 5   

```{r}
err<-mean(round(predict(model,dtest))!=getinfo(dtest,'label'))
print(paste("test-error=", err))
```


### Solution    Exercise 6   

```{r,eval=F}
importance.matrix <- xgb.importance(model = model, 
                                    feature_names = colnames(X))
xgb.plot.importance(importance.matrix)
```

## Importance plot

```{r,echo=F}
importance.matrix <- xgb.importance(model = model, 
                                feature_names = colnames(X))
xgb.plot.importance(importance.matrix)
```


##  Solution   Exercise 7  

```{r}
model_watchlist <- xgb.train(data = dtrain,
                      max_depth = 2,nrounds = 100,
                      objective = "binary:logistic",
                      watchlist = list(train=dtrain, 
                                       test=dtest))
```



##  Solution  Exercise 8   

```{r,eval=F}
model_auc<-xgb.train(data = dtrain,max_depth = 2,
        nrounds = 100,objective = "binary:logistic",
        watchlist = list(train=dtrain,test=dtest),
        eval_metric = 'auc',eval_metric = 'logloss')
```

```{r,eval=F,echo=F}
save(model_auc,file="../data/model_auc.RData")
```

```{r,echo=F}
load("../data/model_auc.RData")
```

## Output `model_auc`

```{r}
model_auc
```


## Solution  Exercise 9   

```{r,eval=F}
library(tidyverse)
model_auc$evaluation_log %>%
  gather(metric, value, -iter) %>%
  separate(metric, c('set','metric')) %>%
  ggplot(aes(iter, value, color = set)) +
  geom_line() +
  facet_grid(metric~.)
```

## Evaluation plot

```{r,echo=F}
library(tidyverse)
model_auc$evaluation_log %>%
  gather(metric, value, -iter) %>%
  separate(metric, c('set','metric')) %>%
  ggplot(aes(iter, value, color = set)) +
  geom_line() +
  facet_grid(metric~.)
```


##  Solution Exercise 10  

```{r,eval=F}
model_eta<-xgb.train(data=dtrain,max_depth = 2,eta = 0.05,
          nrounds = 100,objective = "binary:logistic",
          watchlist = list(train=dtrain, test=dtest),
          eval_metric = 'auc',eval_metric = 'logloss')
```

```{r,echo=F,eval=F}
save(model_eta,file="../data/model_eta.RData")
```

```{r,echo=F}
load("../data/model_eta.RData")
```

## Output of model eta

```{r}
model_eta
```

<!--
https://www.r-exercises.com/2017/10/01/extremely-boost-your-machine-learning-exercises-part-2/

https://www.r-exercises.com/2017/09/24/extremely-boost-your-machine-learning-solutions-part-1/
-->