---
title: 'Visualizing dataset to apply machine learning'
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output: 
  beamer_presentation: 
    theme: Dresden
    colortheme: dolphin
    fig_height: 3
    fig_width: 5
    fig_caption: no
    fonttheme: structuresmallcapsserif
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,message = F,warning=F)
```

## Exercises: Visualizing dataset to apply machine learning

- Exercise based on [r-exercises - visualizing for ml](https://www.r-exercises.com/2017/09/08/visualizing-dataset-to-apply-machine-learning-exercises/)


### Exercise 1

Create a variable “x” and attach to it the input attributes of the “iris” dataset. HINT: Use columns 1 to 4.

### Exercise 2

Create a variable “y” and attach to it the output attribute of the “iris” dataset. HINT: Use column 5.

### Exercise 3

Create a whisker plot (boxplot) for the variable of the first column of the “iris” dataset. HINT: Use `boxplot()`.

## Exercises

### Exercise 4

Now create a whisker plot for each one of the four input variables of the “iris” dataset in one image. HINT: Use par().

### Exercise 5

Create a barplot to breakdown your output attribute. HINT: Use plot().

### Exercise 6

Create a scatterplot matrix of the “iris” dataset using the “x” and “y” variables. HINT: Use featurePlot().

### Exercise 7

Create a scatterplot matrix with ellipses around each separated group. HINT: Use plot="ellipse".

## Exercises

### Exercise 8

Create box and whisker plots of each input variable again, but this time broken down into separated plots for each class. HINT: Use plot="box".

### Exercise 9

Create a list named “scales” that includes the “x” and “y” variables and set relation to “free” for both of them. HINT: Use `list()`

### Exercise 10

Create a density plot matrix for each attribute by class value. HINT: Use `featurePlot()`.


## [Solutions](https://www.r-exercises.com/2017/09/08/visualizing-dataset-to-apply-machine-learning-exercises-solutions/)

###   Solution Exercise 1

<!--
# install.packages("caret")
-->

```{r}
library(caret)
data(iris)
validation <- createDataPartition(iris$Species, p=0.80, 
                                  list=FALSE)
validation20 <- iris[-validation,]
iris <- iris[validation,]
x <- iris[,1:4]
```

###   Solution Exercise 2 

```{r}
library(caret)
y <- iris[,5]
```


##   Solution Exercise 3 

```{r}
library(caret)
boxplot(x[,1], main=names(iris)[1])
```


##    Solution Exercise 4   

```{r}
library(caret)
par(mfrow=c(1,4))
for(i in 1:4) {
  boxplot(x[,i], main=names(iris)[i])
}
```


##    Solution Exercise 5   

```{r}
library(caret)
plot(y)
```


##    Solutions - Visualizing ML

### Solution Exercise 6   

```{r}
library(caret)
featurePlot(x=x, y=y)
```


##    Solution Exercise 7   

```{r}
# install.packages("ellipse")
library(ellipse)
library(caret)
featurePlot(x=x, y=y,plot="ellipse")
```


##    Solutions - Visualizing ML  

### Solution Exercise 8   

```{r}
library(caret)
featurePlot(x=x, y=y, plot="box")
```


##    Solutions - Visualizing ML    

### Solution Exercise 9   

```{r}
library(caret)
scales <- list(x=list(relation="free"), y=list(relation="free"))
```



###    Solution Exercise 10  

```{r,eval=F}
library(caret)
scales <- list(x=list(relation="free"), y=list(relation="free"))
featurePlot(x=x, y=y, plot="density", scales=scales)
```

##    Solution Exercise 10  

```{r,echo=F}
library(caret)
scales <- list(x=list(relation="free"), y=list(relation="free"))
featurePlot(x=x, y=y, plot="density", scales=scales)
```
