---
title: "ML Exercises - neural networks"
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output:
  beamer_presentation: 
    colortheme: dolphin
    fig_height: 3
    fig_width: 5
    fig_caption: no
    fonttheme: structuresmallcapsserif
    highlight: haddock
    theme: Dresden
  pdf_document: 
    keep_tex: yes
    toc: yes
  slidy_presentation: 
    css: mycss.css
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [Exercise neural networks (I)](https://www.r-exercises.com/2017/06/08/neural-networks-exercises-part-1/)



<!--
- Neural network have become a corner stone of machine learning in the last decade.
- Created in the late 1940s with the intention to create computer programs who mimics the way neurons process information


- Those kinds of algorithm have long been believe to be only an academic curiosity, deprived of practical use since they require a lot of processing power and other machine learning algorithm outperform them. However since the mid 2000s, the creation of new neural network types and techniques, couple with the increase availability of fast computers made the neural network a powerful tool that every data analysts or programmer must know.
-->

### Create example data

<!--
Creating the data set on which we want to do a simple regression. 
-->

1) Set the seed to 42, draw 200 numbers from a unified random distribution between -10 and 10 and store them in a vector named `x`. 
2) Create a vector named `y` containing the value of `sin(x)`. 

<!--
Neural network are a lot more flexible than most regression algorithms and can fit complex function with ease. The biggest challenge is to find the appropriate network function appropriate to the situation.
-->

<!--
A network function is made of three components: the network of neurons, the weight of each connection between neuron and the activation function of each neuron.

3)  For this example, we’ll use a feed-forward neural network and the logistic activation which are the defaults for the package `nnet`. We take one number as input of our neural network and we want one number as the output so the size of the input and output layer are both of one. For the hidden layer, we’ll start with three neurons. It’s good practice to randomize the initial weights, so 
-->

### Randomize initial weights

3) Create a vector of 10 random values, picked in the interval [-1,1].

## Solution neuralnet exercises (I)

### Create example data

```{r}
set.seed(42)
x<-runif(200, -10, 10) #1)
y<-sin(x) #2)
```

### Randomize the initial weights

```{r}
weight<-runif(10, -1, 1) #3)
```


## Exercise neural networks (II)

### Split the dataset

4) Neural networks have a strong tendency of overfitting data, they become really good at describing the relationship between the values in the data set, but are not effective with data that wasn’t used to train your model. As a consequence, we need to cross-validate our model. Set the seed to 42, then create a training set containing 75% of the values in your initial data set and a test set containing the rest of your data.

### Create a first model 

5) Load the `nnet` package and use the function `nnet` to create a model. Pass your weights via the `Wts` argument and set `maxit=50`. 
<!--
We fit a function which can have   multiple possible output values. 
-->
Set `linout=T` and take some time to look at the structure of your model.


## Solution neuralnet exercises (II) 


### Split the dataset

```{r}
set.seed(42) #4)
index<-sample(1:length(x),round(0.75*length(x)),replace=FALSE)
reg.train<-data.frame(X=x[index],Y=y[index])
reg.test<-data.frame(X=x[-index],Y=y[-index])
```

### Create a first model 

```{r}
library(nnet)
set.seed(42)
reg.model.1<-nnet(reg.train$X,reg.train$Y,size=3,
                  maxit=50,Wts=weight,linout=TRUE)
```

<!--
## # weights:  10
## initial  value 103.169943 
## iter  10 value 70.636986
## iter  20 value 69.759785
## iter  30 value 63.215384
## iter  40 value 45.634297
## iter  50 value 39.876476
## final  value 39.876476 
## stopped after 50 iterations
-->

## Solution neuralnet exercises (III)

```{r}
str(reg.model.1)
```

<!--
## List of 15
##  $ n            : num [1:3] 1 3 1
##  $ nunits       : int 6
##  $ nconn        : num [1:7] 0 0 0 2 4 6 10
##  $ conn         : num [1:10] 0 1 0 1 0 1 0 2 3 4
##  $ nsunits      : num 5
##  $ decay        : num 0
##  $ entropy      : logi FALSE
##  $ softmax      : logi FALSE
##  $ censored     : logi FALSE
##  $ value        : num 39.9
##  $ wts          : num [1:10] -7.503 2.202 3.004 -0.806 -4.69 ...
##  $ convergence  : int 1
##  $ fitted.values: num [1:150, 1] -0.196 0.568 -0.353 -0.205 -0.161 ...
##   ..- attr(*, "dimnames")=List of 2
##   .. ..$ : NULL
##   .. ..$ : NULL
##  $ residuals    : num [1:150, 1] 0.692 0.3079 -0.0398 0.4262 -0.7024 ...
##   ..- attr(*, "dimnames")=List of 2
##   .. ..$ : NULL
##   .. ..$ : NULL
##  $ call         : language nnet.default(x = reg.train$X, y = reg.train$Y, size = 3, Wts = weight,      linout = TRUE, maxit = 50)
##  - attr(*, "class")= chr "nnet"
-->


## Exercise neural networks (III)

### Prediction

6) Predict the output for the test set and compute the RMSE of your predictions. Plot the function sin(x) and then plot your predictions.

### Repeat with different parameters

7) The number of neurons in the hidden layer, and the number of hidden layer used, has a great influence on the effectiveness of your model. Repeat the exercises four to six, but this time use a hidden layer with seven neurons and initiate randomly 22 weights.


## Solution neuralnet exercises (IV)

### Prediction

```{r}
# 6)
predict.model.1<-predict(reg.model.1,data.frame(X=reg.test$X))
str(predict.model.1)
```

<!--
##  num [1:50, 1] -0.201 0.184 -0.873 -0.981 0.598 ...
##  - attr(*, "dimnames")=List of 2
##   ..$ : NULL
##   ..$ : NULL
-->

```{r}
# 6)
rmse.reg<-sqrt(sum((reg.test$Y-predict.model.1)^2))
rmse.reg
```

<!--
## [1] 3.41651
-->

## Solution neuralnet exercises (V)

```{r}
# 6)
plot(sin, -10, 10)
points(reg.test$X,predict.model.1)
```

<!--
plot of chunk NN_part1
-->

## Solution neuralnet exercises (V)
<!--
Exercise 6
-->

### Repeat with different parameters

```{r}
# 7)
set.seed(42)
reg.model.2<-nnet(reg.train$X,reg.train$Y,size=7,
                  maxit=50,Wts=runif(22, -1, 1),linout=TRUE)
```

## Solution neuralnet exercises (VI)

```{r}
# 7)
str(reg.model.2)
```


## Solution neuralnet exercises (VII)

```{r}
# 7)
predict.model.2<-predict(reg.model.2,data.frame(X=reg.test$X))
str(predict.model.2)
```

```{r}
# 7)
rmse.reg<-sqrt(sum((reg.test$Y-predict.model.2)^2))
rmse.reg
```

<!--
## [1] 2.188407
-->

## Solution neuralnet exercises (VIII)

```{r}
# 7)
plot(sin, -10, 10)
points(reg.test$X,predict.model.2)
```



## Exercise neural networks (IV)

### Normilze

8) Now let us use neural networks to solve a classification problem, so let’s load the iris data set! It is good practice to normalize your input data to uniformize the behavior of your model over different range of value and have a faster training. Normalize each factor so that they have a mean of zero and a standard deviation of 1, then create your train and test set.

### Create model with `nnet`

9) Use the `nnet()` and use a hidden layer of ten neurons to create your model. We want to fit a function which have a finite amount of value as output. To do so, set the linout argument to true. Look at the structure of your model. With classification problem, the output is usually a factor that is coded as multiple dummy variables, instead of a single numeric value. As a consequence, the output layer have as one less neuron than the number of levels of the output factor.


<!--
plot of chunk NN_part1
-->

## Solution Exercise 8

```{r}
data<-iris

scale.data<-data.frame(lapply(data[,1:4], function(x) scale(x)))
scale.data$Species<-data$Species
index<-sample(1:nrow(scale.data),round(0.75*nrow(scale.data)),
              replace=FALSE)
clust.train<-scale.data[index,]
clust.test<-scale.data[-index,]
```


## Solution Exercise 9

```{r}
set.seed(42)
clust.model<-nnet(Species~.,size=10,Wts=runif(83, -1, 1),
                  data=clust.train)
```

<!--
## # weights:  83
## initial  value 187.294915 
## iter  10 value 10.386561
## iter  20 value 5.337510
## iter  30 value 2.311922
## iter  40 value 1.426508
## iter  50 value 1.387440
## iter  60 value 1.386324
## final  value 1.386294 
## converged
-->

## Exercise neural networks (V)

### prediction and confusion table

10) Make prediction with the values of the test set.
 
11) Create the confusion table of your prediction and compute the accuracy of the model.



## Solution Exercise 10 and 11

### Prediction

```{r}
# 10)
predict.model.clust<-predict(clust.model,clust.test[,1:4],
                             type="class")
```

### Confusion matrix

```{r}
# 11)
(Table<-table(clust.test$Species ,predict.model.clust))
```

<!--
##             predict.model.clust
##              setosa versicolor virginica
##   setosa         16          0         0
##   versicolor      0          9         0
##   virginica       0          0        13
-->

```{r}
(accuracy<-sum(diag(Table))/sum(Table))
```


<!--
https://www.r-exercises.com/2017/06/21/neural-networks-solutions-part-3/

https://www.r-exercises.com/?s=neural+networks
-->


