---
title: "Machine Learning - Solution/Exercises"
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output:
  beamer_presentation: 
    colortheme: dolphin
    fig_height: 3
    fig_width: 5
    fig_caption: no
    fonttheme: structuresmallcapsserif
    highlight: haddock
    theme: Dresden
  pdf_document: 
    keep_tex: yes
    toc: yes
  slidy_presentation: 
    css: mycss.css
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=T,warning = F,message = F)
```

## [Exercise - `rpart` Kyphosis](https://www.r-exercises.com/2016/12/13/recursive-partitioning-and-regression-trees-exercises/)

### Consider the Kyphosis data frame

<!--
(type `help("kyphosis")` for more details)
-->

1) Which variables are in the `kyphosis` dataset
2) Build a tree to classify Kyphosis from Age, Number and Start.

### Consider the tree build above.

3) Which variables are used to explain Kyphosis presence?
4) How many observations contain the terminal nodes.

### Consider the Kyphosis data frame.

5) Build a tree using the first 60 observations of kyphosis.
6) Predict the kyphosis presence for the other 21 observations.
7) Which is the misclassification rate (prediction error)

## [The dataset kyphosis](https://www.r-exercises.com/2016/12/13/recursive-partitioning-and-regression-trees-solutions/)

### The dataset contains (1):

- Kyphosis: a factor with levels absent present indicating if a kyphosis (a type of deformation) was present after the operation.
- Age: in months.
- Number: the number of vertebrae involved.
- Start: the number of the first (topmost) vertebra operated on.

## Build the tree (2)

```{r}
library('rpart')
TREE <- rpart(Kyphosis ~ Age + Number + Start, 
              data=kyphosis,method="class")
TREE
```

## Plot the result

```{r}
plot(TREE)
text(TREE,use.n=T)
```

## Answers

3) Which variables are used to explain Kyphosis presence?

- The variables are Start and Age

4) How many observations contain the terminal nodes.

- *denotes terminal nodes. The nodes have 29, 12, 14, 7 and 19 observations

## Consider the Kyphosis data frame.

5) Build a tree using the first 60 observations of kyphosis.

```{r}
TREE <- rpart(Kyphosis ~ Age + Number + Start, 
              data=kyphosis[1:60,],method="class")
```


6) Predict the kyphosis presence for the other 21 observations.

```{r}
PR <- predict(TREE,kyphosis[61:81,],type='class')
```


7) Which is the misclassification rate (prediction error) 

```{r}
test=kyphosis$Kyphosis[61:81]
table(PR,test)
```

```{r}
rate <- 100*length(which(PR!=test))/length(PR)
cat('the misclassification rate is:',rate)
```

## Exercise `rpart` -  `iris`

### Consider the `iris` data frame

1) Build a tree to classify Species from the other variables.
2) Plot the trees, add nodes information.

### Consider the tree build before

3) Prune the the using median complexity parameter (cp) associated to the tree.
4) Plot in the same window, the pruned and the original tree.
5) In which terminal nodes is clasified each oobservations of `iris`?
6) Which Specie has a flower of `Petal.Length` greater than 2.45 and `Petal.Width` less than 1.75.

## Solution - `rpart` -  `iris` (I)

1) Build a tree to classify Species from the other variables.

```{r}
TREE2=rpart(Species ~ ., data=iris,method="class")
TREE2
```

## Solution - `rpart` -  `iris` (II)


2) Plot the trees, add nodes information.

```{r}
plot(TREE2)
text(TREE2,use.n=T)
```

## Solution - `rpart` -  `iris` (III)


3) Prune the the using median complexity parameter (cp) associated to the tree.

```{r}
TP <- prune(TREE2,cp=median(TREE2$cptable[,'CP']))
```

4) Plot in the same window, the pruned and the original tree.

```{r,fig.height=3}
par(mfrow=c(1,2))
plot(TREE2);text(TREE2,use.n=T)
plot(TP);text(TP,use.n=T)
```


## Solution - `rpart` -  `iris` (IV)

5) In which terminal nodes is clasified each observations of iris?

```{r}
TREE2$where
```

## Solution - `rpart` -  `iris` (V)


6) Which species has a flower of `Petal.Length` greater than 2.45 and `Petal.Width` less than 1.75.

```{r}
print('versicolor')
```

