---
title: "ML exercises - basics R"
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output:
  slidy_presentation: 
    css: mycss.css
    keep_md: yes
  pdf_document: 
    keep_tex: yes
    toc: yes
  beamer_presentation: 
    colortheme: dolphin
    fig_height: 3
    fig_width: 5
    fig_caption: no
    fonttheme: structuresmallcapsserif
    highlight: haddock
    theme: Dresden
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,warning=F,message=F)
```


## Exercise: Find R-packages

Go to https://cran.r-project.org/ and search for packages that can be used:

1) to reduce overfitting
2) for regression trees
3) for gradient boosting
4) for neural networks
5) for clustering

## Solution: Find R-packages

```{r,eval=F}
install.packages("glmnet") #1)
install.packages("rpart") #2)
install.packages("gbm") #3)
install.packages("neuralnet") #4)
install.packages("kknn") #5)
```

## Exercise: load built-in data

### Load the the built-in dataset `swiss` 
<!--
and answer the following questions:
-->
1) How many observations and variables are available?
2) What is the scale level of the variables?

### Interactive data table 

3) Create an interactive data table

## Solution: load built-in data

```{r}
# 1)
data(swiss) 
dim(swiss) 
str(swiss) 
```

```{r,eval=F}
# 2)
DT::datatable(swiss)
```

## [Exercise](https://www.datacamp.com/community/tutorials/pipe-r-tutorial): random numbers

```{r,echo=F}
x <- c(0.109, 0.359, 0.63, 0.996, 0.515, 0.142, 0.017, 
       0.829, 0.907)
x <- runif(8)
```

1) Draw 8 random numbers from the uniform distribution and save them in a vector `x`
2) Compute the logarithm of `x`, return suitably lagged and iterated differences, 
3) compute the exponential function and round the result

```{r,echo=F}
round(exp(diff(log(x))), 1)
```

## Solution: random numbers

```{r,echo=F}
x <- runif(8) #1)
round(exp(diff(log(x))), 1) #2) and 3)
```



<!--
Exercise cross validation
https://www.r-exercises.com/2017/09/15/applying-machine-learning-algorithms-exercises/

Applying machine learning algorithms – exercises: solutions

https://www.r-exercises.com/2017/11/28/machine-learning-with-h2o-part-3-exercises/
-->
