---
title: "Introduction to R"
author: "Jan-Philipp Kolb"
date: "26 April 2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction round

- Where are you from? What are you studying/working?
- What are your expectations of this course?
- Where do you think you can use Machine Learning in the future?

## Preliminaries

- This topic is huge - we concentrate on presenting the application in R
- Usually we have big differences in knowledge and abilities of the participants - please tell, if it is too fast or slow.
- We have many [**exercises**](http://web.math.ku.dk/~helle/R-intro/exercises.pdf) because at the end you can only learn on your own
- We have many [**examples**](https://www.showmeshiny.com/) - try them out 
- If there are questions - always ask
- R is more fun together - ask your neighbor

## Why R is a good choice ...

- ... because it is an [**open source language**](https://stackoverflow.com/questions/1546583/what-is-the-definition-of-an-open-source-programming-language)
- ... outstanding graphs - [**graphics**](http://matthewlincoln.net/2014/12/20/adjacency-matrix-plots-with-r-and-ggplot2.html), [**graphics**](https://www.r-bloggers.com/3d-plots-with-ggplot2-and-plotly /), [**graphics**](https://procomun.wordpress.com/2011/03/18/splomr/)
- ... relates to other languages - [**R can be used in combination with other programs**](https://github.com/Japhilko/RInterfaces) - e.g. [**data linking**](https://github.com/Japhilko/RInterfaces/blob/master/slides/Datenimport.md)
- ...R can be used [**for automation**](https://cran.r-project.org/web/packages/MplusAutomation/index.html)
 - ... Vast Community - [**you can use the intelligence of other people ;-)**](https://www.r-bloggers.com/)
- ...

- Because of the large comunity
- New statistical methodologies are implemented quite fast 
- Because R can be combined with other programs like Postgresql or Python

## Constraints

<!--
Why R might not be the best choice in every situation
-->

### Newer modules in Python

- Machine learning is a field that changes rapidly.  
- Some new tools are first developed in Python. 
- The package reticulate offers the possibility to use these modules from an R environment. 
- Good news - Python is also Open Source

### Big Data 

- Especially if you work with web data, you quickly have to deal with large amounts of data. 
- Therefore one must fall back on databases, which can be used in combination with R. 

## Content of this part

- Introduction to programming in R 

### what is relevant for this course.

- How to import data?
- What to do with missing values?
- Parallelization




## Import data 

```{r,eval=F}
?read.csv
?read.csv2
```

```{r,eval=F}
?scan
```

## The R-package `data.table`

### Get an overview

```{r}
data(airquality)
head(airquality)
```

### Overview with `data.table`

```{r}
library(data.table)
airq <- data.table(airquality)
airq
```



## How to deal with missing values

```{r,eval=F}
?na.omit
```

```{r}
airq
```

```{r}
na.omit(airq)
```


## [Time measurement](https://www.r-bloggers.com/5-ways-to-measure-running-time-of-r-code/)

```{r}
start_time <- Sys.time()
ab <- runif(10000000)
end_time <- Sys.time()

end_time - start_time
```


## How many cores are available


```{r}
library(doParallel)
detectCores()
```

## Make cluster

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```

```{r}
start_time <- Sys.time()
ab <- runif(10000000)
end_time <- Sys.time()

end_time - start_time
```

```{r}
stopCluster(cl)
```


```{r,eval=F}
?parallel::makeCluster
```

