---
title: "Clustering"
author: "Jan-Philipp Kolb and Alexander Murray-Watters"
date: "18 Januar 2019"
output: beamer_presentation
---

```{r setupClustering, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




## Resources


```{r,echo=F, eval=FALSE}
slides_path <- getwd()
git_path <- gsub("slides","",slides_path)
if (Sys.info()$nodename=="MAC14077"){
  git_path <- "D:/Daten/GitHub/machine_learning/"
  slides_path <- paste0(git_path,"/slides")
}
```


- [Package `kknn`](https://cran.r-project.org/web/packages/kknn/kknn.pdf)

```{r,eval=F}
install.packages("kknn")
```

```{r}
library("kknn")
```


## [Geographic clustering of UK cities](https://www.r-bloggers.com/geographic-clustering-of-uk-cities/)

Animated example: 
https://towardsdatascience.com/the-5-clustering-algorithms-data-scientists-need-to-know-a36d136ef68


## Exercise: Kmeans

Apply kmeans to to the `iris` dataset with 2, 3, and 4
clusters. Produce three scatter plots, with the points colored
according to cluster assignment.



## hdbscan

A fairly new alternative to kmeans, hdbscan does not require you to
specify the number of categories to be assigned. It only requires a
decision as to the minimum number of points needed to be included in a
cluster. This minimum number acts as a smoothing parameter (such as a
density bandwidth parameter or a histograms bin/bar width), with lower
values finding more clusters. Other advantages of hdbscan include .

```{r, eval=FALSE}
install.packages("dbscan")
```



```{r}
library(ggplot2)
library(dplyr)
library(maps)
library(dbscan)

## Example where kmeans finds only 1 cluster.
two.clust.eg <- rbind(matrix(rnorm(1000, sd = 0.8), ncol=2),
                      matrix(rnorm(100, mean = 120, sd = 0.12), ncol = 2))

clust <- kmeans(two.clust.eg, centers=2)

plot(x, col = clust$cluster)
##     points(cl$centers, col = 1:2, pch = 8, cex = 2)


```

```{r}





data(moons)

## Running HDBscan with the minimum number of points set to 5.
res <- dbscan::hdbscan(moons, minPts = 3)

plot(moons, col = res$cluster + 1, main="R implementation")
```



## Exercise: Apply kmeans to the moons dataset and compare the results. 
-- Be sure to try different numbers of centers.


## Exercise: Apply hdbscan to the moons dataset with different minimums for the number of points. 

## Exercise: Apply both kmeans and hdbscan to the `ChickWeight` dataset's "weight" "Time" variables, and see how well you can get each to perform.




## Exercise: Apply hdbscan to the moons dataset with different minimums for the number of points. 
```{r, eval=FALSE, echo=FALSE}
## kmeans
plot(ChickWeight[,1:2], col=kmeans(ChickWeight[,1:2], centers=4)$centers)

## hdbscan, minPts=10
plot(ChickWeight[,1:2], col=dbscan::hdbscan(ChickWeight[,1:2], minPts=10)$cluster)

## Diet cat. for comparison.
plot(ChickWeight[,1:2], col=ChickWeight$Diet)

## Chick cat. for comparison.
plot(ChickWeight[,1:2], col=ChickWeight$Chick)


```


```{r, eval=FALSE}
load(paste0(git_path,"/data/osmsa_PLZ_14.RData"))
```



## [US Census Data](https://elitedatascience.com/datasets)

- [US Census Data (Clustering)](https://archive.ics.uci.edu/ml/datasets/US+Census+Data+%281990%29) – Clustering based on demographics is a tried and true way to perform market research and segmentation.



## Links

- [Using clusterlab to benchmark clustering algorithms](https://www.r-bloggers.com/using-clusterlab-to-benchmark-clustering-algorithms/)
