---
title: "Clustering"
author: "Jan-Philipp Kolb"
date: "18 Januar 2019"
output: beamer_presentation
---

```{r setupClustering, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




## Resources


```{r,echo=F, eval=FALSE}
slides_path <- getwd()
git_path <- gsub("slides","",slides_path)
if (Sys.info()$nodename=="MAC14077"){
  git_path <- "D:/Daten/GitHub/machine_learning/"
  slides_path <- paste0(git_path,"/slides")
}
```


- [Package `kknn`](https://cran.r-project.org/web/packages/kknn/kknn.pdf)

```{r,eval=F}
install.packages("kknn")
```

```{r}
library("kknn")
```


## [Geographic clustering of UK cities](https://www.r-bloggers.com/geographic-clustering-of-uk-cities/)



## hdbscan

A fairly new alternative to kmeans, hdbscan does not require you to
specify the number of categories to be assigned. It only requires a
decision as to the minimum number of points needed to be included in a
cluster. This minimum number acts as a smoothing paramter (such as a
density bandwidth parameter or a histograms bin/bar width), with lower
values finding more clusters. Other advantages of hdbscan include .

```{r, eval=FALSE}
install.packages("dbscan")
```



```{r}
library(ggplot2)
library(dplyr)
library(maps)
library(dbscan)


data(moons)

## Running HDBscan with the minimum number of points set to 5.
res <- dbscan::hdbscan(moons, minPts = 3)

plot(moons, col = res$cluster + 1, main="R implementation")
```






```{r, eval=FALSE}
load(paste0(git_path,"/data/osmsa_PLZ_14.RData"))
```



## [US Census Data](https://elitedatascience.com/datasets)

- [US Census Data (Clustering)](https://archive.ics.uci.edu/ml/datasets/US+Census+Data+%281990%29) – Clustering based on demographics is a tried and true way to perform market research and segmentation.



## Links

- [Using clusterlab to benchmark clustering algorithms](https://www.r-bloggers.com/using-clusterlab-to-benchmark-clustering-algorithms/)
